Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-01-03 lqg-strings.tex
Added the recent quantization progress to main file

\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} G_{\lambda \nu} + \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} G_{\lambda \nu} = \\ = \frac{1}{2} F^{-} \sqrt{- g} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \partial _{\lambda} G_{\mu \nu} ,
\end{multline}
where $F^{-} = 1 - k \Delta / 2 (- g)$. Using conformal symmetry $g^{a b} = (\phi (x))^{-1} \eta ^{a b}$ and assuming flat background $G_{\mu \nu} = \eta _{\mu \nu}$, eq(\ref{eqn:X-eom-inverse-area}) simplify to
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \partial _{a} \ln (F^{-}) \partial _{b} X^{\mu} = 0 .
\end{equation}

@@

Equation(\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal function $\phi (x)$ appearing in $F^{-}$. By using the ansatz in eq(\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal function to fix the second term as $- m^2 X^{\mu}$, with $- m^2 = p^2 - E^2$ on-shell, and this in turn implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

@@

At first glance, our recent analysis of the Polyakov string with inverse metric correction seems to not be related to what we found in the analysis of the Nambu-Goto string. However, as pointed out in \cite{Grimshaw2010Homogenization}, wave equations with a variable speed (the result of the Nambu-Goto analysis) can be turned into Klein-Gordon equations (what we just found). While they're work deals only with space-varying wave speed, it is straight forward to generalise the procedure to also include time dependence, but ironically enough the calculations lead to the conclusion that the wave speed $v = 1 / F^{+}$ is actually time independent.

@@

In the following, we proceed to quantize the Klein-Gordon string via canonical covariant quantization, in a manner much similar to those found in \cite{Tong2009Lectures} and \cite{Green_Schwarz_Witten_2012}. To start, we calculate the momentum canonical to the embbeding fields $X^{\mu}$, which reads
\begin{multline} \label{eqn:canonical-momentum}
	\Pi _{\mu} (\tau , \sigma) = F^{-} \frac{i}{(4 \pi \alpha ')^2} \sum _{p} \left( - \alpha ^{\nu} _{p} \e ^{i (- E_{p} \tau + p \sigma)} + \\
	+ (\alpha ^{\nu} _{- p})^{*} \e ^{i (E_{p} \tau + p \sigma)} \right) \eta _{\nu \mu} ,
\end{multline}
and then we proceed to turn the equal-time Poisson bracket relations
\begin{equation} \label{eqn:Poisson-brackets-1}
	\left\{ X^{\mu} (\tau , \sigma) , \Pi ^{\nu} (\tau , \sigma ') \right\} = \delta (\sigma - \sigma ') \eta ^{\mu \nu}
\end{equation}
\begin{equation} \label{eqn:Poisson-brackets-2}
	\left\{ X^{\mu} (\tau , \sigma) , X^{\nu} (\tau , \sigma ') \right\} = 0 = \left\{ \Pi ^{\mu} (\tau , \sigma) , \Pi ^{\nu} (\tau , \sigma ') \right\}
\end{equation}
into equal-time commutation relations between the now operators $\Hat{X}^{\mu}$ and $\Hat{\Pi}^{\nu}$:
\begin{equation}
	\Big[ \Hat{X}^{\mu} (\tau , \sigma) , \Hat{\Pi}^{\nu} (\tau , \sigma ') \Big] \equiv i \delta (\sigma - \sigma ') \eta ^{\mu \nu}
\end{equation}

>>>>

\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} G_{\lambda \nu} + \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} G_{\lambda \nu} = \\ = \frac{1}{2} F^{-} \sqrt{- g} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \partial _{\lambda} G_{\mu \nu} ,
\end{multline}
where $F^{-} = 1 - k \Delta / 2 (- g)$. Using conformal gauge $g^{a b} = (\phi (x))^{-1} \eta ^{a b}$ and assuming flat background $G_{\mu \nu} = \eta _{\mu \nu}$, eq(\ref{eqn:X-eom-inverse-area}) simplify to
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \partial _{a} \ln (F^{-}) \partial _{b} X^{\mu} = 0 .
\end{equation}

@@

Equation(\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal factor $\phi (x)$ appearing in $F^{-}$. By using the ansatz in eq(\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal function to fix the second term as $- \mu ^2 X^{\mu}$, with $- \mu ^2 = p^2 - E^2$ on-shell, and this in turn implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

@@

At first glance, our recent analysis of the Polyakov string with inverse metric correction seems to not be related to what we found in the analysis of the Nambu-Goto string. However, as pointed out in \cite{Grimshaw2010Homogenization}, wave equations with a variable speed (the result of the Nambu-Goto analysis) can be turned into Klein-Gordon equations (what we just found). While their work deals only with space-varying wave speed, it is straightforward to generalise the procedure to also include time dependence.

@@

In the following, we proceed to quantize the Klein-Gordon string via canonical covariant quantization, in a manner much similar to those found in \cite{Tong2009Lectures} and \cite{Green_Schwarz_Witten_2012}. To start, we calculate the momentum canonical to the embbeding fields $X^{\mu}$, which reads
\begin{multline} \label{eqn:canonical-momentum}
	\Pi _{\mu} (\tau , \sigma) = F^{-} \frac{i}{(4 \pi \alpha ')^2} \sum _{p} \left( - \alpha ^{\nu} _{p} \e ^{i (- E_{p} \tau + p \sigma)} + \\
	+ (\alpha ^{\nu} _{- p})^{*} \e ^{i (E_{p} \tau + p \sigma)} \right) \eta _{\nu \mu} ,
\end{multline}
and then we proceed to turn the equal-time Poisson bracket relations
\begin{equation} \label{eqn:Poisson-brackets-1}
	\left\{ X^{\mu} (\tau , \sigma) , \Pi ^{\nu} (\tau , \sigma ') \right\} = \delta (\sigma - \sigma ') \eta ^{\mu \nu}
\end{equation}
\begin{equation} \label{eqn:Poisson-brackets-2}
	\left\{ X^{\mu} (\tau , \sigma) , X^{\nu} (\tau , \sigma ') \right\} = 0 = \left\{ \Pi ^{\mu} (\tau , \sigma) , \Pi ^{\nu} (\tau , \sigma ') \right\}
\end{equation}
into equal-time commutation relations between the now operators $\Hat{X}^{\mu}$ and $\Hat{\Pi}^{\nu}$:
\begin{equation} \label{eqn:canonical-commutator-1}
	\Big[ \Hat{X}^{\mu} (\tau , \sigma) , \Hat{\Pi}^{\nu} (\tau , \sigma ') \Big] \equiv i \delta (\sigma - \sigma ') \eta ^{\mu \nu}
\end{equation}
\begin{equation} \label{eqn:canonical-commutator-2}
	\Big[ \Hat{X}^{\mu} (\tau , \sigma) , \Hat{X}^{\nu} (\tau , \sigma ') \Big] = 0 = \Big[ \Hat{\Pi}^{\mu} (\tau , \sigma) , \Hat{\Pi}^{\nu} (\tau , \sigma ') \Big] .
\end{equation}
From these relations, we get commutation relations for the vibrational modes $\Hat{\alpha}^{\mu} _{p}$ and $(\Hat{\alpha}^{\nu} _{- k})^{\dagger}$:
\begin{equation} \label{eqn:non-vanishing-commutator}
	\Big[ \Hat{\alpha}^{\mu} _{p} , (\Hat{\alpha}^{\nu} _{- k})^{\dagger} \Big] = \eta ^{\mu \nu} \delta _{p , - k}
\end{equation}
\begin{equation} \label{eqn:vanishing-commutators}
	\Big[ \Hat{\alpha}^{\mu} _{p} , \Hat{\alpha}^{\nu} _{k} \Big] = 0 = \Big[ (\Hat{\alpha}^{\mu} _{- p})^{\dagger} , (\Hat{\alpha}^{\nu} _{- k})^{\dagger} \Big] ,
\end{equation}
which are just harmonic oscillator relations with
\begin{align}
	\Hat{a}^{\mu} _{p}             & = \Hat{\alpha}^{\mu} _{p} \label{eqn:annihilation-operator}                                       \\
	(\Hat{a}^{\mu} _{p})^{\dagger} & = (\Hat{\alpha}^{\mu} _{p})^{\dagger} , \ \mathrm{for} \ p \neq 0 . \label{eqn:creation-operator}
\end{align}
For $p = 0$ we have the center of mass momentum operator $\Hat{p}^{\mu}$. As usual, we define a vacuum state for the string to obey
\begin{equation} \label{eqn:annihilation-vacuum}
	\Hat{a}^{\mu} _{p} \ket{0} = 0 , \ \mathrm{for} \ p \neq 0 ,
\end{equation}
and for $p = 0$
\begin{equation} \label{eqn:momentum-vacuum}
	\Hat{p}^{\mu} \ket{0} = p^{\mu} \ket{0} .
\end{equation}
Our Fock space is built from the vacuum state $\ket{0}$ by operating with a sequence of creation operators
\begin{multline} \label{eqn:general-state}
	\ket{\psi} = ((a^{\mu _{1}} _{1})^{\dagger})^{n_{\mu _1}} ((a^{\mu _2} _{2})^{\dagger})^{n_{\mu _2}} ... \times \\
	\times ((a^{\nu _1} _{1})^{\dagger})^{n_{\nu _1}} ((a^{\mu _2} _{2})^{\dagger})^{n_{\nu _2}} ... \ket{0} .
\end{multline}

As usual, by the appearance of the Minkowski metric in the non-zero commutator
\begin{equation} \label{eqn:ghosts}
	\Big[ \Hat{a}^{\mu} _{p} , (\Hat{a}^{\nu} _{k})^{\dagger} \Big] = \eta ^{\mu \nu} \delta _{p k} ,
\end{equation}
we have ghosts in the theory which must be dealt with. When translating the constraints as quantum operators, we require that they have vanishing matrix elements when sandwiched by physical states $\psi$ and $\phi$:
\begin{equation} \label{eqn:L-constraint-quantum-1}
	\bra{\phi} \Hat{L}_{p , k} \ket{\psi} = 0 \, \,
\end{equation}
\begin{equation} \label{eqn:L-constraint-quantum-2}
	\bra{\phi} \Hat{\widetilde{L}}_{p , k} \ket{\psi} = 0 .
\end{equation}
$\Hat{L}_{p , k}$ has no ambiguities since it's composed of only annihilation operators, however, we have an ambiguity for $\Hat{\widetilde{L}}_{p , 0}$. To resolve this issue, we pick
\begin{equation} \label{eqn:L-normal-ordering}
	\Hat{\widetilde{L}}_{p , 0} = \frac{- p}{E_{- p}} (\Hat{\alpha}^{\mu} _{p})^{\dagger} \Hat{\alpha}^{\nu} _{p} \eta _{\mu \nu}
\end{equation}
with the annihilation operator to the right, which manifests in the imposition of the constraint as
\begin{equation} \label{eqn:shifted-L-constraint}
	\bra{\phi} \Hat{\widetilde{L}}_{p , 0} - c_{p} \ket{\psi} = 0 ,
\end{equation}
for some sequence $c_{p}$. The reason for it to be a sequence of $p$ and not a constant like in the regular String Theory literature is that in the usual theory, we would have a sum over $p$ and thus it would end up being a constant, but in our analysis we have this extra freedom, which we'll now see ends up being important. Classically, we had eq(\ref{eqn:string-rest-mass}) for the string rest mass, which we can rewrite as
\begin{equation} \label{eqn:alternative-classical-rest-mass}
	M^2 = - \frac{1}{(4 \pi (\alpha ')^2)^2} \lim _{n \rightarrow 0} \left( - \frac{E_{p}}{p} \widetilde{L}_{p , 0} \right) ,
\end{equation}
from where we can now see that the string mass spectrum will get shifted by
\begin{equation} \label{eqn:shifted-mass-spectrum}
	\Hat{M}^2 = - \frac{1}{(4 \pi (\alpha ')^2)^2} \left( (\Hat{\alpha} ^{\mu} _{0})^{\dagger} \Hat{\alpha}^{\nu} _{0} \eta _{\mu \nu} + \lim _{p \rightarrow 0} \frac{E_{p}}{p} c_{p} \right) .
\end{equation}
Since $E_{0} = \mu$, it seems like the natural candidate for $c_{p}$ would be some function $f (\mu , p)$ such that
\begin{equation} \label{eqn:cp-limit}
	\lim _{p \rightarrow 0} \frac{c_{p}}{p} = \mu
\end{equation}
like $c_{p} = \mu p$, $c_{p} = \sin (\mu p)$, $c_{p} = \tan (\mu p)$, etc, making it so the mass spectrum of the string gets shifted by the \textquotedblleft mass" term of the Klein-Gordon equation (\ref{eqn:X-Klein-Gordon-eqn}), $\mu ^2$.

The \textquotedblleft Virasoro-Klein-Gordon" algebra our constraint modes satisfy is given by
\begin{equation} \label{eqn:L-commutators-1}
	\Big[ \Hat{L}_{n , m} , \Hat{L}_{p , k} \Big] = 0
\end{equation}
\begin{multline} \label{eqn:L-commutators-2}
	\Big[ \Hat{L}_{n , m} , \Hat{\widetilde{L}}_{p , k} \Big] = \frac{m - n}{E_{m - n}} \Big( \Hat{L}_{n + k - p , n} \delta _{n - m , k} + \\
	+ \Hat{L}_{n - m + k - p , n - m} \delta _{n , k} \Big)
\end{multline}
\begin{multline} \label{eqn:L-commutators-3}
	\Big[ \Hat{\widetilde{L}}_{n , m} , \Hat{\widetilde{L}}_{p , k} \Big] = \frac{(m - n)}{E_{m - n}} \frac{(k - p)}{E_{k - p}} \times \\
	\times \Bigg( \frac{E_{- n}}{(- n)} \Hat{\widetilde{L}}_{k - p - n , k - p} \delta _{n - m , k} + \\
	+ \frac{E_{- k}}{(- k)} \Hat{\widetilde{L}}_{n - m - k , n - m} \delta _{n , k - p} \Bigg)
\end{multline}
